---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  comment = "#>", 
  fig.path = "man/figures/README-", 
  fig.align="center", 
  fig.width = 8
)
```

# rjd3filters

rjd3filters is an R package on linear filters for real-time trend-cycle estimates.
It allows to create symmetric and asymmetric moving averages with:

- local polynomial filters, as defined by Proietti and Luati (2008);

- the FST approach  of Grun-Rehomme, Guggemos, and Ladiray (2018), based on the optimization of the three criteria Fidelity, Smoothness and Timeliness;

- the Reproducing Kernel Hilbert Space (RKHS) of Dagum and Bianconcini (2008).

Some quality criteria defined by Wildi and McElroy (2019) can also be computed.

## Installation

rjd3filters relies on the [rJava](https://CRAN.R-project.org/package=rJava) package.

Running rjd3 packages requires **Java 17 or higher**. How to set up such a configuration in R is explained [here](https://jdemetra-new-documentation.netlify.app/#Rconfig) 

To get the current stable version (from the latest release):

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("rjdverse/rjd3toolkit@*release")
remotes::install_github("rjdverse/rjd3filters@*release")
```

To get the current development version from GitHub:

```{r, eval = FALSE}
# install.packages("remotes")
# Install development version from GitHub
remotes::install_github("rjdverse/rjd3toolkit")
remotes::install_github("rjdverse/rjd3filters")
```


## Basic example

In this example we use the same symmetric moving average (Henderson), but we use three different methods to compute asymmetric filters. As a consequence, the filtered time series is the same, except at the boundaries.

```{r plot-global, message = FALSE}
library("rjd3filters")

y <- window(retailsa$AllOtherGenMerchandiseStores, start = 2000)
musgrave <- lp_filter(horizon = 6, kernel = "Henderson", endpoints = "LC")

# we put a large weight on the timeliness criteria
fst_notimeliness_filter <- lapply(0:6, fst_filter, 
                                  lags = 6, smoothness.weight = 1/1000, 
                                  timeliness.weight = 1-1/1000, pdegree =2)
fst_notimeliness <- finite_filters(sfilter = fst_notimeliness_filter[[7]], 
                                   rfilters = fst_notimeliness_filter[-7], 
                                   first_to_last = TRUE)
# RKHS filters minimizing timeliness
rkhs_timeliness <- rkhs_filter(horizon = 6, asymmetricCriterion = "Timeliness")

trend_musgrave <- filter(y, musgrave)
trend_fst <- filter(y, fst_notimeliness)
trend_rkhs <- filter(y, rkhs_timeliness)
plot(ts.union(y, trend_musgrave, trend_fst, trend_rkhs), plot.type = "single", 
     col = c("black", "orange", "lightblue", "red"), 
     main = "Filtered time series", ylab=NULL)
legend("topleft", legend = c("y", "Musgrave", "FST", "RKHS"), 
       col= c("black", "orange", "lightblue", "red"), lty = 1)
```

The last estimates can also be analysed with the `implicit_forecast` function that retreive the implicit forecasts corresponding to the asymmetric filters (i.e., the forecasts needed to have the same end-points estimates but using the symmetric filter).

```{r plot-forecast}
f_musgrave <- implicit_forecast(y, musgrave)
f_fst <- implicit_forecast(y, fst_notimeliness)
f_rkhs <- implicit_forecast(y, rkhs_timeliness)

plot(window(y, start = 2007), 
     xlim = c(2007, 2012), ylim = c(3600, 4600), 
     main = "Last estimates and implicit forecast", ylab=NULL)
lines(trend_musgrave, 
      col = "orange")
lines(trend_fst, 
      col = "lightblue")
lines(trend_rkhs, 
      col = "red")
lines(ts(c(tail(y, 1), f_musgrave), frequency = frequency(y), start = end(y)), 
      col = "orange", lty = 2)
lines(ts(c(tail(y, 1), f_fst), frequency = frequency(y), start = end(y)), 
      col = "lightblue", lty = 2)
lines(ts(c(tail(y, 1), f_rkhs), frequency = frequency(y), start = end(y)), 
      col = "red", lty = 2)
legend("topleft", legend = c("y", "Musgrave", "FST", "RKHS", "Forecasts"), 
       col= c("black", "orange", "lightblue", "red", "black"), 
       lty = c(1, 1, 1, 1, 2))
```

The real-time estimates (when no future points are available) can also be compared:

```{r plot-q0, eval = TRUE}
trend_henderson<- filter(y, musgrave[, "q=6"])
trend_musgrave_q0 <- filter(y, musgrave[, "q=0"])
trend_fst_q0 <- filter(y, fst_notimeliness[, "q=0"])
trend_rkhs_q0 <- filter(y, rkhs_timeliness[, "q=0"])
plot(window(ts.union(y, trend_musgrave_q0, trend_fst_q0, trend_rkhs_q0), 
            start = 2007), 
     plot.type = "single", 
     col = c("black", "orange", "lightblue", "red"), 
     main = "Real time estimates of the trend", ylab=NULL)
legend("topleft", legend = c("y", "Musgrave", "FST", "RKHS"), 
       col= c("black", "orange", "lightblue", "red"), lty = 1)
```

### Comparison of the filters

Different quality criteria from Grun-Rehomme *et al* (2018) and Wildi and McElroy(2019) can also be computed with the function `diagnostic_matrix()`:

```{r diagnostic-table, eval = TRUE}
q_0_coefs <- list(Musgrave = musgrave[, "q=0"], 
                  fst_notimeliness = fst_notimeliness[, "q=0"], 
                  rkhs_timeliness = rkhs_timeliness[, "q=0"])

sapply(X = q_0_coefs, 
       FUN = diagnostic_matrix, 
       lags = 6, 
       sweights = musgrave[, "q=6"])
```

The filters can also be compared by plotting there coefficients (`plot_coef`), gain function (`plot_gain`) and phase function (`plot_phase`):

```{r, diagnostic-plots, eval = FALSE}
def.par <- par(no.readonly = TRUE)
par(mai = c(0.3, 0.3, 0.2, 0))
layout(matrix(c(1, 1, 2, 3), 2, 2, byrow = TRUE))

plot_coef(fst_notimeliness, q = 0, col = "lightblue")
plot_coef(musgrave, q = 0, add = TRUE, col = "orange")
plot_coef(rkhs_timeliness, q = 0, add = TRUE, col = "red")
legend("topleft", legend = c("Musgrave", "FST", "RKHS"), 
       col= c("orange", "lightblue", "red"), lty = 1)

plot_gain(fst_notimeliness, q = 0, col = "lightblue")
plot_gain(musgrave, q = 0, col = "orange", add = TRUE)
plot_gain(rkhs_timeliness, q = 0, add = TRUE, col = "red")
legend("topright", legend = c("Musgrave", "FST", "RKHS"), 
       col= c("orange", "lightblue", "red"), lty = 1)

plot_phase(fst_notimeliness, q = 0, col = "lightblue")
plot_phase(musgrave, q = 0, col = "orange", add = TRUE)
plot_phase(rkhs_timeliness, q = 0, add = TRUE, col = "red")
legend("topright", legend = c("Musgrave", "FST", "RKHS"), 
       col= c("orange", "lightblue", "red"), lty = 1)
par(def.par)
```

```{r, diagnostic-plots, eval = TRUE, echo=FALSE}
```

Confidence intervals can also be computed with the `confint_filter` function:

```{r, confint-plot, eval = TRUE}
confint <- confint_filter(y, musgrave)

plot(confint, plot.type = "single",
     col = c("red", "black", "black"),
     lty = c(1, 2, 2), xlab = NULL, ylab = NULL)
lines(y, col = "grey")
legend("topleft", legend = c("y", "Smoothed", "CI (95%)"), 
       col= c("grey", "red", "black"), lty = c(1, 1, 2))
```

### Manipulate moving averages

You can also create and manipulate moving averages with the class `moving_average`.
In the next examples we show how to create the M2X12 moving average, the first moving average used to extract the trend-cycle in X-11, and the M3X3 moving average, applied to each months to extract seasonal component.


```{r, mm-plots, eval = TRUE}
e1 <- moving_average(rep(1, 12), lags = -6)
e1 <- e1/sum(e1)
e2 <- moving_average(rep(1/12, 12), lags = -5)
M2X12 <- (e1 + e2)/2
coef(M2X12)
M3 <- moving_average(rep(1/3, 3), lags = -1)
M3X3 <- M3 * M3
# M3X3 moving average applied to each month
M3X3
M3X3_seasonal <- to_seasonal(M3X3, 12)
# M3X3_seasonal moving average applied to the global series
M3X3_seasonal

def.par <- par(no.readonly = TRUE)
par(mai = c(0.5, 0.8, 0.3, 0))
layout(matrix(c(1, 2), nrow = 1))
plot_gain(M3X3, main = "M3X3 applied to each month")
plot_gain(M3X3_seasonal, main = "M3X3 applied to the global series")
par(def.par)

# To apply the moving average
t <- y * M2X12
si <- y - t
s <- si * M3X3_seasonal
# or equivalently:
s_mm <- M3X3_seasonal * (1 - M2X12)
s <- y * s_mm
```

### Manipulate finite filters

`finite_filters` object are a combination of a central filter (used for the final estimates) and different asymmetric filters used for intermediate estimates at the beginning/end of the series when the central filter cannot be applied.

```{r}
musgrave 
musgrave * M3X3
```


## Bibliography

Dagum, Estela Bee and Silvia Bianconcini (2008). “The Henderson Smoother in Reproducing Kernel Hilbert Space”. In: *Journal of Business & Economic Statistics 26*, pp. 536–545. URL: [https://ideas.repec.org/a/bes/jnlbes/v26y2008p536-545.html](https://ideas.repec.org/a/bes/jnlbes/v26y2008p536-545.html).

Grun-Rehomme, Michel, Fabien Guggemos, and Dominique Ladiray (2018). “Asymmetric Moving Averages Minimizing Phase Shift”. In: *Handbook on Seasonal Adjustment*. URL: https://ec.europa.eu/eurostat/web/products-manuals-and-guidelines/-/KS-GQ-18-001.

Proietti, Tommaso and Alessandra Luati (Dec. 2008). “Real time estimation in local polynomial regression, with application to trend-cycle analysis”. In: *Ann. Appl. Stat.* 2.4, pp. 1523–1553. URL: https://doi.org/10.1214/08-AOAS195.

Wildi, Marc and Tucker McElroy (2019). “The trilemma between accuracy, timeliness and smoothness in real-time signal extraction”. In: *International Journal of Forecasting* 35.3, pp. 1072–1084. URL: [https://EconPapers.repec.org/RePEc:eee:intfor:v<wbr>:35:y:2019:i:3:p:1072-1084](https://EconPapers.repec.org/RePEc:eee:intfor:v:35:y:2019:i:3:p:1072-1084).


## Package Maintenance and contributing

Any contribution is welcome and should be done through pull requests and/or issues.
pull requests should include **updated tests** and **updated documentation**. If functionality is changed, docstrings should be added or updated.


## Licensing

The code of this project is licensed under the [European Union Public Licence (EUPL)](https://joinup.ec.europa.eu/collection/eupl/eupl-text-eupl-12).
